### 2020-05-13 paper.tex
read poblete--steered MD using a particular RMSD for RNA

\documentclass{article}
\title{Reading Notes}

\begin{document}
\title{}

today I got this repo set up. Next time, I'll reskim these papers:

MARTINI\cite{Marrink2007}

VOTCA\cite{Ruhle2009}

TraPPE\cite{Maerzke2011}

and read this: \cite{Chakraborty2018a}

got busy with fancy-ing up the GIXStapose repo, helping rachel, office hours, and atom-typing with gaff-foyer with ryan and chris. Briefly skimmed the chakraborty paper-- they consider all possible mappings that a cg compound can have in order to systematically pick a cging scheme.
OK Chakraborty notes:
-exclude cg mapping schemes to those that maintain symmetry
-Zhang and Voth designed CG mapping operators by preserving "essential dynamics"
-Cao and Voth worked on centering functions COM vs center of charge. They found center of change was better for preserving intermolecular structure.
-Wagner generalize centering functions to arbitrary thermodynamic quantities
-Marrink grouped four heavy atoms and showed that this simple consistent scheme works.
CG models are validated by preservation of rdfs and velocity autocorrelation functions (vacfs)
mapping operator tree

Today I'm reading the VOTCA\cite{Ruhle2009} paper with the goal of answering these questions:
how do they choose a chromophore?

how do they calculate the transfer integral?

how do they do backmapping?

other notes:
CG methods--IBI, force matching, inverse MC. This paper goes over details of these and how they're implemented in votca. This wasn't really what I wanted to read about so I skimmed.
They use ESPResSo++ for quantum calculations
OK, this isn't the right paper to read about CT stuff or backmapping. Let's look at some of these [http://www.votca.org/citations/using-votca](http://www.votca.org/citations/using-votca)
Also I'm noticing in this list of papers--Marrink (cited by chakraborty) is here and there is a paper mentioning the importance of side-chains for charge-mobility! \cite{May2011}
Next time read about tie-chains \cite{May2011} and/or charge mobility \cite{Lukyanov2010}
Lukyanov--sounds a lot like morphct, gaussian density of states? I thought their (votca's) method was like morphct but now I'm not sure... It seems like they're using a simplification and not doing dft.

https://pubs.acs.org/doi/10.1021/ja104380c - 
rubrene record charge carrier $15 \frac{cm^{2}}{(V s)}$ single crystal by vapor deposition. typical for thin-film is $1 {cm^{2}}{(V s)}$

good discussion of some of the assumption of marcus theory. 
Drude model (electrons bounce off stationary nuclei like a pinball machine) works for perfect crystals at low temp.
At RT it doesn't work. Marcus theory assumes nuclear dynamics much slower (Borne-Oppenheimer) and electronic coupling is weak (what does this mean exactly?). Static disorder is based on the electronic DOS and the hopping rate between localized states. (As opposed to what?)
They start with single crystal structures from xray data. They calculate reorganization energies with b3lyp 6-311G(d,p) and transfer integrals with ZINDO. (check their SI)
How are they able to calculate direction-resolved transfer integrals?? very cool. What exactly do the transfer integral graphs mean? (Figure 2) Had a helpful chat with Chris about this and it helped me to think through the figures. The y-axis on the transfer integral is counts and the x-axis is energies--it makes sense that movement of an electron in different directions would require different energies and/or be more likely. Still not sure how to calculate it though.

\cite{Vehoff2010a}
rubrene record charge carrier $15 \frac{cm^{2}}{V s}$ single crystal by vapor deposition. typical for thin-film is $1 \frac{cm^{2}}{V s}$

Today I'm digging around in more of the votca stuff to try and figure out how to run kmc there. They use ORCA too! Also their tutorials and documentation REALLY SUCK. [https://github.com/votca/xtp-tutorials](https://github.com/votca/xtp-tutorials) in particular I'm looking at LAMMPS/KMC\_Thiophene.
OK. either I am dumb or that tutorial sucks. I can tell that they're using orca and the inputs/oputputs the use but there's no explanation of what they're doing. It looks like they calculate the reorganization energy, but they don't define a chromophore--which makes sense for thiophene but what about when your molecule is ginormous?
\cite{Ruhle2011b} discusses charge transport in amorphous films of [tris-(8-hydroxyquinoline)aluminum](https://en.wikipedia.org/wiki/Tris(8-hydroxyquinolinato)aluminium). Oh Figure 1 is nice--it is a workflow of what can be done in votca.
11A talks about how to parameterize new contributions to a force-field using first-principles.
11B choosing a "chromophore"-- rigid fragments vs conjugated segments. they broke apart the chromophore in a polymer like p3ht using the dihedral angle--although they validated this choice with other methods. Interesting-- they replace the chromophores with those which have been geometry optimized using DFT. then they make a neighbor list
I'll come back to this paper later. I at least was able to answer some of my questions about choosing a chromophore.
11C talks about calculating the transfer integral for each pair and why ZINDO works and when it breaks down.

rereading the perspective paper \cite{Jankowski2019} looking for things to read next:
Braun 2018 foundations in molecular simulation
https://doi.org/10.33011/ livecoms.1.1.5957
74-89 are relevant to coarse-graining
92-95 start here for opvs

Today I'm thinking about some of the big picture stuff related to my proposal:
Science-wise I am thinking of my proposal like this: Humans use energy and burning fossil fuels to get that energy really sucks. rising CO2 concentration in the atmosphere contributes to global warming. Sustainable, low-carbon footprint energy generation is vital to protect the precious climate which sustains all life. Energy generation using OPV devices is one form of alternative energy. But there are a lot of potential OPV molecules out there! So how can we develop inexpensive, fast methods for screening this ZOO of compounds? Well we can use modelling techniques and a computer. Computational simulations of organic photovoltaics suffer from a problem of length scales: OPV morphologies can have bulk features on the scale of 10s of nanometers (BHJ, exciton dissociation length, pi-stacking) which affect the morphology, yet still we are interested in the electronic properties of this system--a fundementally quantum mechanical property--taking place on the angstrom scale where with current computers we can only realistically look at up to 100's of atoms. <so we break apart the problem MD to get a morphology, Zindo to calculate energy splitting of neighbors-->transfer integral, KMC to estimate CT, talk about results of evan/matty/votca> What am I going to add? more robust/automated coarse-graining and backmapping, calculation of reorganization energy, chromophore detection.
Things I care about: The environment--I like working on a project that theoretically could help find the next molecule for OPV technology. Sometimes it seems very far removed, but having sustainable technology as an ultimate goal matters to me. I also care a lot about help make science more accessible to everyone: I find myself really excited about creating tools which help to visualize the unseeable and to communicate complex ideas in a simple understandable way. I'm not sure this has anything to do with anything, but it's definitely a developing passion.

I'm also reading this paper today https://arxiv.org/pdf/2003.02031.pdf which is a bout TRUE simulations.
notes:
GUI's hurt reproducibility, MS Excel hides the formula's and order of calculation (p 4)
poorly documented code and raw data files dont help reproducibility (looking at you votca) (p4)
*I want to include start to finish notebooks with any paper I publish...* Also docker containers with my exact software stack.
I love the examples in this paper :) It feels a lot like our perspective paper! I'm skimming them but I love that notebooks are included in the SI.
They qualify the GUI's are bad statement saying that GUIs are only bad *if they hide the workflow details.*

Matt Thompson got a job at OpenFF, one of the tools they list as desirable for applicants is knowledge of psi4:
https://pubs.acs.org/doi/abs/10.1021/acs.jctc.8b00286

read official papers for mosdef. check github.

Reading \cite{Hanson2016} also my mendeley overwriting my files because it thinks they are duplicates or something? It keeps losing the pdf even though I've uploaded it. :(
SMILES - simplified molecular-input line-entry system
SMARTS - smiles arbitrary target specification
around since 1980s
alternatives SLN, MQL, CSRML.
This paper is about specifically Jmol smiles so I'm skipping to a different one.
\cite{Weininger1988} First paper about smiles- how do we unambiguously represent a chemical structure? historically CAS, graph theory, IUPAC.
human-readable and machine readable
hydrogens, single bond don't need to be specified. designed for organic molecules.

\cite{Smith2018} reading about psi4: free and open source with python api that uses blas and c under the hood. has jupyter notebooks tutorials. This paper reads like our perspective paper in tone :hearteyes: 2.2 -2.3 kinda lost me. I should look at their jupyter notebooks.

Today searched "coarse grain backmapping" in the library and got the following papers:
\cite{Alessandri2017b}

Coarse graining in OPV bulk heterojunctions! p3ht and pcbm, solvent evaporation. morphologies need large interfacial area between donor and acceptor and relatively high crystallinity and connection to the electrode.How can we investigate the processing conditions that achieve the highest performance?
Analysis of OPVs include AFM, SEM, TEM - provide info about interpenetration (how the two phases are dispersed.) Electron tomography can study 3d morphology., but has limited contrast in organic material due to low scattering. GIXD, GISAXS quantifies the degree of crystallinity but averages over the structure. 
CG allows us to look at these systems but CG models are currently not transferrable meaning that for each new system a new CG model must be developed which is slow. Simulated annealing is often used but a more relevant to experiment method is simulated evaporation.
This paper used CG model like Lee and Pao (read this paper) but also based on finer model like martini. Martini does backmapping. (backwards.py) 
They use explicit solvent (CB/benzene) and they remove molecules randomly as the simulation progresses. Longer drying time results in larger crystallites/ fewer D-A contacts. P3ht shows backbone (pi-pi) stacking at a distance of 4.5 A in the CG morphology. They use backwards.py to do their backmapping. They use 3 beads to represent thiophene in p3ht with no explicitly anisotropic interaction. Carillo (ref 26) used anisotropic addition onto very simplified model (thiophene is single bead) of lee (ref 22). (read these) The model used here thiophene is a triangle and they think this helps it stay planar and stack.
Figure 5 is beautiful.

\cite{Haxton2015a}
CG beads with orientation. They use the word "stiffest" to describe degrees of freedom--they mean bond stretching, angle bending, and torsion within rings. They suggest there are two separate ways to improve CG: better mapping between AA and CG or better CG interactions. The seconds can be helped with IBI, entropy minimization... First one is what the authors will focus on.
Point-site method maps groups of atoms onto a bead--bead potential is spherically symmetric, but some models use neighboring sites to introduce some directionality. e.g., Alemani et al created a protein model where the dipole interaction was a function of the alpha-carbon angle. Here they add orientation to the point-site model. This increases the degrees of freedom for each site from 3 to 6, but they model bonds and angles as stiff and dihedral angles (except rings and double bonds) as soft. They also map about 8 atomns to a site. They use backmapping from ref 64--not backwards.py. 
They simplify the ideal interactions of the CG model (many-body problem) to pai potentials. They show how to calculate to error in an observable calculated on backmapped structure. C code used for backmapping in SI.
mapping ratio - atomic DOF / CG DOF. PRIMO on average is 3.0, UNRES, MARTINI, MS-CG are 8.2, 7.1, and 6.0. I think this paper aims to improve on PRIMO by adding directionality? they measure accuracy with rmsd and efficiency with mapping ratio.

\cite{Wassenar2014b}

This paper is about backwards.py. The authors are from university of groeningen where GROMACS is made. however this tool is generic and not specific to gromacs. it uses geometric projection and position restrained MD relaxation--same as what Matty did. can map between anything just needs a mapping.
Good overview of CG methods:
Rzepiela - randomly place atomistic beads on small spheres around CG beads. Requires many steps, specific to one verison of GROMACS, but can be made specific with correspondence between atoms and beads.
Brocos - lipid micelles. placing atoms on a trace through the CG bead and adding peripheral atoms at offset. 
A good CG model should yield correct structures and be simple quick and flexible. Should be generic (ff and system type) and require little information. Then information provided should be the CG coordinates, the atomistic topology and the correspondence mapping between the atomistic and CG particles.
They go from MARTINI-CG to GROMOS 54a7 UA to CHARMM36 AA.
Some matrix mapping that I don't totally understand. Confusing between A the matrix and A the atom position. :/
Fig 5 is neat! c: shows a helical protein at each step of the CG->AA process. their procedure appears to improve upon the simulated anealing of rzepiela. 
* look at backwards.py with the intent of better understanding the Matrix transformation.

\cite{Peng2019a}

biomolecules. Backmapping based on Bayesian inference and restrained MD. cgmap.py -- they tell us the names of their codes but don't provide them?? They try multiple CG methods: Martini (twice with different initial structure), UNRES, ENM. Martini had the lowest RMSD but it is also the finest grain method. They have an ultra coarse method that only use 5 beads for a small protein; it gets impressive results (RMSD 0.47 angstrom) for how sparse it is! But it is not sufficiently accurate to predict structure-- other RMSD > 10angstrom. their method is more expensive. I don't quite understand how they get some of the probabilities for the Bayesian part, but my takeaway from this paper is that the logrestraints work a lot better for them and Bayesian inference, although expensive, allows them to skip the AA MD simulation at the end?

\cite{Rolland2020a}

This is the paper we reviewed in lab meeting! CG -> backmap -> mobility. AA MD can reproduce XRD patterns pi-pi stacking, formation of lamellar structure, change in morphology at high oxidation levels, mobility dependence on chain-length (* looks into ref 26 and compare to evan's tie chain paper)
CG allows us to look at larger lengthscales and with backmapping we can restore atomistic details, but is there a cost? ref 44 model HOTT molecules as ellipsoids with Gay Berne interactions (I'm thinking of this as anisotropic--i.e., directionally dependent so good for ellipsoids--Lennard Jones. Often used in liquid crystal systems.) 
They use the same backmapping as us ref 13. Many CG papers verify by XRD buyt they claim same XRD can result in very different mobility. ref 26. They use Wassenar's backward.py script. -- He is from the university which makes GROMACS
CG overestimates density--backmapping fixes it. XRD of backmapped AA match AA. pipi stacking is corrected and lamellar structure present. Mobility also good. They explain the differences between the AA and CG mobilities saying that the AA one is smaller so it's a size effect due to over estimated percolation threshhold--i.e. the connectivity is overestimated because the periodic volume is small enough that the periodic image has an effect.

\cite{Poblete2018a}
RNA - backmapping steered MDaimed at minimizing a particular RMSD (ERMSD) between atomic and target structure. target structure were found to be recovered by this backmapping except in the case where there were unpaired bases. folding pathway dependent on parameters used in ERMSD.
need AA strutures to validate with experiment (x-ray or neutron scattering) backmapping usually involves placing atomistic coordinates relative to the beads and then minimizing the energy of the structure using steepest descent or short MD runs (could use position restraint). Atomistic can even be mapped onto the whole trajectory. 
This work describes the backmapping of the SPlit and conQueR (SPQR) method. Minimizes the ERMSD distance between solvated AA strand of the CG structure (RNA). ERMSD : origin is set to CoM of C2 C4 and C6 atoms of each base (see ref 23), then for each pair of bases, rij is the center of base i in the coord system of base j. In order to account for anisotropy, the scaling for the z-direction is different (does this mean the strand is oriented to z in some way??) This metric shows the difference between arrangement of nucleobases of two RNA structures. Backbone and sugar interactions are unrestrained. The paper will show results on the backbone conformation. They use steered MD with ERMSD as the steering parameter. (How does steered MD work?) Look like they use and external pulling force which depend on ERMSD.

\bibliography{library}
\bibliographystyle{ieeetr}
\end{document}