### 2020-08-17 hexane.tex
update hexane images, tex file to generate small pdf

\documentclass[8pt]{extarticle}

\usepackage[paperwidth=90mm]{geometry}

%PUT YOUR MACROS HERE
\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,language=xml,aboveskip=3mm,belowskip=3mm,showstringspaces=false,columns=flexible,basicstyle={\small\ttfamily},numbers=none,numberstyle=\tiny\color{gray},keywordstyle=\color{blue},commentstyle=\color{dkgreen},stringstyle=\color{mauve},breaklines=true,breakatwhitespace=true,tabsize=4}

\begin{document}

\lstinputlisting{hexane.xml}

\end{document}